Hybrid MILP to efficiently and accuratly \\ solve hard DNN verification instances

### 2024-09-17 ICLR25/Paper_noProof.tex
tables

In this paper, we study the MILP encoding of ReLU-DNNs, provide new deep insights in the LP relaxation, which allows us to carefully craft a {\em partial MILP} solution which only considers few neurons as integer variables, the rest using the LP relaxation. Compared with previous attempts, we can reduce the number of integer variables by around 4 times while maintaining the same level of accuracy. This gives rise to a very efficient yet accurate algorithm. Implemented in {\em Hybrid MILP (hMILP)}, calling first $\alpha,\beta$-CROWN with a short time-out to solve easier instances, and then partial MILP for those for which $\alpha,\beta$-CROWN fails, produces a very accurate yet efficient verifier, reducing tremendously the gap of undecided instances to $\approx 10\%$, while keeping a reasonable runtime ($<500s$ on average per instance).
\end{abstract}

@@

\begin{table}
	\centering
	\begin{tabular}{||l|c|c||c|c|c||}
		\hline
		Network & Acc & Upper  & $\alpha,\beta$-Crown& $\alpha,\beta$-Crown & $\alpha,\beta$-Crown \\ 
		Perturbation &   & Bound & TO=10s & TO=30s & TO=2000s\\ \hline
		MNIST 6$\times$100 & 99\% & 90\% & 33\% & 35\% & 40\%   \\
		$\epsilon = 0.026$ &  &  & 6.9s &  18.9s &  1026.1s  \\  \hline
		MNIST 6$\times$200 & 99\%  & 96\%  & 46\%  & 49\%  & 50\%   \\ 
		$\epsilon = 0.015$ & &  & 6.5s &  16.6s &  930.3s  \\  \hline
		MNIST 9$\times$100 & 97\%  & 86\%  & 23\%  & 28\%  & 28\%   \\
		$\epsilon = 0.026$ &  &  & 7.2s &  20.1s &  930.3s  \\  \hline
		MNIST 9$\times$200 & 97\%  & 91\%  & 35\%  & 36\%  & 37\%   \\ 
		$\epsilon = 0.015$ & &  & 6.8s &  18.2s &  1083.1s  \\  \hline
		MNIST 6$\times$500 & 100\%  & 94\%  & 41\%  & 43\%  & 44\%   \\ 
		$\epsilon = 0.035$ & &  & 6.4s &  16.4s &  1002.5s  \\  \hline
		CIFAR CNN-B-adv & 78\%  & 62\%  &  34\% & 40\%  & 42\%   \\
		$\epsilon = 2/255$&  &  & 4.3s & 8.7s & 373s  \\ \hline \hline
		CIFAR ResNet & 29\%  & 25\%  & 25\%  & 25\%  & 25\%   \\
		$\epsilon = 2/255$ &  &  & 2s & 2s & 2s  \\ \hline
	\end{tabular}
	\caption{Verification $\%$ of $\alpha,\beta$-Crown run on 7 DNNs (average runtime per image in second). The 6 first DNNs are hard instances, while the last one (ResNet) is an easy instance (it is trained to be easy to verify using Wong, but with a very low accuracy level), provided for reference.}
	\label{table_beta}
\end{table}

@@

\begin{table}
	\centering
\begin{tabular}{||l|c|c||c|c|c||}
	\hline
	Network &  Acc & Upper Bound & Marabou 2.0  & Full MILP & NNenum? \\\hline
	MNIST 6$\times$100 & 99\% & 90\% & 26\% & 30\% & ?\%   \\
	$\epsilon = 0.026$ & &  & 6520s & 7000s &  ?s  \\  \hline
\end{tabular}
\caption{Complete verifiers run on the smallest of the considered DNN. 
Because the instances are hard, most test time-out at the global threshold of 10.000s per image, and no complete verifier could outperform 
$\alpha,\beta$-CROWN (40\%) for this DNN, despite the much larger average runtime. We didnt test on larger DNNs as this small one already took considerable amount of time.}
\label{table_complete}
\end{table}

>>>>

In this paper, we study the MILP encoding of ReLU-DNNs, provide new deep insights in the LP relaxation, which allows us to carefully craft a {\em partial MILP} solution which selects automatically few neurons encoded as integer variables, the rest using the LP relaxation. Compared with previous attempts, we can reduce the number of integer variables by around 4 times while maintaining the same level of accuracy. This gives rise to a very efficient yet accurate algorithm. Implemented in {\em Hybrid MILP (hMILP)}, calling first $\alpha,\beta$-CROWN with a short time-out to solve easier instances, and then partial MILP for those for which $\alpha,\beta$-CROWN fails, produces a very accurate yet efficient verifier, reducing tremendously the gap of undecided instances to $\approx 10\%$, while keeping a reasonable runtime ($<500s$ on average per instance).
\end{abstract}

@@

\begin{table}
	\centering
	\begin{tabular}{||l|c|c||c|c|c||}
		\hline
		Network & Acc & Upper  & $\alpha,\beta$-Crown& $\alpha,\beta$-Crown & $\alpha,\beta$-Crown \\ 
		Perturbation &   & Bound & TO=10s & TO=30s & TO=2000s\\ \hline
		MNIST 6$\times$100 & 99\% & 90\% & 33\% & 35\% & 40\%   \\
		$\epsilon = 0.026$ &  &  & 6.9s &  18.9s &  1026s  \\  \hline
		MNIST 6$\times$200 & 99\%  & 96\%  & 46\%  & 49\%  & 50\%   \\ 
		$\epsilon = 0.015$ & &  & 6.5s &  16.6s &  930s  \\  \hline
		MNIST 9$\times$100 & 97\%  & 86\%  & 23\%  & 28\%  & 28\%   \\
		$\epsilon = 0.026$ &  &  & 7.2s &  20.1s &  930s  \\  \hline
		MNIST 9$\times$200 & 97\%  & 91\%  & 35\%  & 36\%  & 37\%   \\ 
		$\epsilon = 0.015$ & &  & 6.8s &  18.2s &  1083s  \\  \hline
		MNIST 6$\times$500 & 100\%  & 94\%  & 41\%  & 43\%  & 44\%   \\ 
		$\epsilon = 0.035$ & &  & 6.4s &  16.4s &  1003s  \\  \hline
		CIFAR CNN-B-adv & 78\%  & 62\%  &  34\% & 40\%  & 42\%   \\
		$\epsilon = 2/255$&  &  & 4.3s & 8.7s & 373s  \\ \hline \hline
		CIFAR ResNet & 29\%  & 25\%  & 25\%  & 25\%  & 25\%   \\
		$\epsilon = 2/255$ &  &  & 2s & 2s & 2s  \\ \hline
	\end{tabular}
	\caption{Verification $\%$ of $\alpha,\beta$-Crown run on 7 DNNs (average runtime per image in second). The 6 first DNNs are hard instances, while the last one (ResNet) is an easy instance (it is trained to be easy to verify using Wong, but with a very low accuracy level), provided for reference.}
	\label{table_beta}
\end{table}

@@

\begin{table}
	\centering
	\begin{tabular}{||c||c|c|c||}
		\hline
		Network &  Marabou  & & NNenum? Full MILP  \\ \hline
		MNIST 6$\times$100 & 28\% & $\approx 30\%$ & 40 \%    \\ \hline
		$\epsilon = 0.026$ & 6200s &  5588s & 6100s
		  \\  \hline
	\end{tabular}
\caption{Complete verifiers run on the smallest of the considered DNN. 
Because the instances are hard, most test time-out at the global threshold of 10.000s per image, and no complete verifier could outperform 
$\alpha,\beta$-CROWN (40\%, 1026s) for this DNN, despite the much larger average runtime. We didnt test on larger DNNs as this small one already took considerable amount of time.}
\label{table_complete}
\end{table}
